### 2019-03-20 papers/journal/intro.tex
Intro: Additon complete: All Mukund's suggestions incorporated

The next step is membrane deformation, which involves cytosolic coat proteins. 
%
Coat proteins are recruited to the forming vesicle through their interaction with adaptor proteins. 
%
These are naturally curved proteins, and their assembly on membranes leads to membrane deformation.
%
A single kind of coat protein can bind to multiple kinds of adaptors, thereby allowing the packaging of multiple types of cargo within the same vesicle. 
%
There are three major types of coat proteins: Clathrin, COP2 and COP1. These act on different compartments: the clathrin coat operates at multiple compartments, and can interact with multiple adaptors proteins, whereas the COP2 and COP1 coat are only involved in vesicle formation at a single compartment; the ER and the Golgi respectively~\cite{faini2013vesicle}.

Finally, transport vesicles bud out from the source compartment. This process requires ATP and is performed by the cytosolic dynamin proteins, which are recruited to the vesicle neck by the coat proteins~\cite{cocucci2014dynamin}.
%
The vesicle is now ready for transport and subsequent fusion with a target compartment.

@@

Target compartment recognition involves molecular switches called Rabs and  tethering complexes. Distinct compartments of the cell are associated with distinct Rabs. For example, Rab5 is associated early endosomes and  Rab7 with late endosomes\cite{rink2005rab}.  Rabs are master regulators; in their active form Rabs recruit many downstream proteins, including tethering complexes, fusion regulators, motor proteins, etc, to their compartment. ~\cite{muller2018molecular}

Tethering complexes sequester vesicles to their target compartments by binding to the vesicle on one end and the target compartment on the other end, thereby bridging the two before fusion. They are also believed to function as chaperones for assembly of SNARE protein complexes, which enable vesicle fusion. 
A clear mechanism for action of these tethers is missing, but evidence from experiments indicate that tethers bind to vesicles through molecular interactions with vesicle coat proteins, vesicle associated Rabs, SNAREs and SNARE regulators. They can also sense membrane curvature (vesicles are highly curved) ~\cite{baker2016chaperoning}.

The final step in vesicle transport is fusion of vesicle with the target compartment.This is brought about by SNARE proteins. 
There are four kinds of SNAREs: R, Qa, Qb, and Qc; and vesicle fusion requires formation of a SNARE complex containing one of each type. In all known cases, one SNARE protein of the complex is contributed by the vesicle and the other three SNAREs by the target compartment (A few SNARE proteins, such as SNAP25, contain both Qb- and Qc-SNARE motifs)~\cite{yoon2018snare}.

%\somya{schematic of snare regulatory network}
In keeping with their central role in vesicle transport, both the activity and the localization of SNARE proteins are regulated. Notable regulators are Sec17/18 and the SM proteins.
%
Sec17/18 bring about disassembly of SNARE complexes post-fusion in order that these SNAREs may be reused. 
%
Whereas, SM proteins have following multiple modes of SNARE regulation: 
\begin{itemize}
	\item SM proteins can hold
	the Qa-SNAREs in an inhibited state and prevent or postpone their assembly into SNARE
	complexes.
	%
	\item Some SM proteins can act as a template upon which an early SNARE complex intermediate can
	form.
	%
	\item Finally, SM proteins bind the fully assembled
	SNARE complexes to protect them from premature disassembly by Sec17/18~\cite{baker2016chaperoning}.
	%
	\item SM proteins have also been shown to interact with tethers~\cite{yoon2018snare}.
\end{itemize}

@@

\subsection{The model}
\ankit{A template: Requires rewriting and some explanation}
%
\todo{Ankit: The Rothman model has no intuition; Add couple of lines relating it to VTS; Also, relate it to previous subsection. The connect is lacking.} The VTS model for this paper is inspired by~\cite{shukla2017discovering}. 
%
On the timescales of minutes, we have tried to capture important aspect of Rothman-Schekman-Sudhof (RSS) model~\cite{rothman2002machinery} of vesicle traffic system.
%
The following are the basic component and assumptions of our model. 
\begin{enumerate}
\item A cell is a set of compartments exchanging molecules via vesicles of uniform size.
\item On the timescale of minutes, compartments are neither created nor destroyed.
\item Each compartment is in steady state, gain and loss of the molecules is balance\todo{Ankit: Need intuition and also reasoning why this is okay to add this as an assumption}.
\item We assume that molecules are neither created nor destroyed.
\item Molecules move via vesicles of uniform size.
\item Identical vesicles have identical target compartments.
\item Fusion of vesicles to compartments is driven by specific SNARE pairing.
\item The identities of compartments and vesicles is encoded in their molecules, and identical vesicles fuse with identical target compartments.
\item The activity of a SNARE can be regulated by other molecules present on the same compartment
or vesicle.
\item An active SNARE pair is necessary and sufficient for fusion. 
\end{enumerate}

@@

\textbf{Steady state:} 
The vesicle transfer will change the molecular composition (distinct molecule count) on both the source compartment and the target compartment. 
%
In our model, we assume that the cell is in steady state. As a further simplification, we also assume that the incoming and outgoing flux is balanced for each of the molecular types at each compartment. 
Indeed, it is well accepted that on the scale of minutes and hours the molecular composition remains the same. 
%
In this paper, we refer to the steady state of the cell (refered as \textit{homeostasis} in biological literature) as the stability condition over the annotated graph.

@@

%
%Properties of the VTS would be hindered by the same reason. 
%3-connected graphs and all possible variations of SNARE regulation rules. 
%
%The number of graphs of specified connectivity grows exponentially with the number of nodes: Table~\ref{tab:threec} shows how many 3-edge-connected graphs~\cite{a052448-oeis} exist (without parallel or self edges) as node number N increases.
\subsection{Graph connectivity is an interesting and accessible property of the VTS.} 
\ankit{Requires rewriting and better explanation}
%\textbf{Properties of a VTS that satisfies all cell-biological constraints:} 
Interactions between the molecules of the VTS together produce a functional vesicle transport network, in which molecules flow between intracellular compartments using specified paths.
%
The molecules of the VTS, their interactions and modes of regulation have been worked out to a large extent through cell-biological experiments. 
%
But we do not know if this information is complete. In order to assess how completely we know the VTS, we need easily readable benchmarks.

@@

%\ankit{We need not specify Synthesis here. Only connectivity use next subsection to build this.}
%\somya{But this is the para where we explain why connectivity is interesting an biological property}
%\ankit{Yes; Both of our remark looks correct!}
The vesicle transport graphs available today for eukaryotic organisms are incomplete themselves. 
%
In this situation, we use our result as a tool to predict missing edges which satisfy the graph connectivity constraint. 
%
These predictions are intended to be used as guides for future experiments. 
%
A mismatch between our predictions and experiments would indicate that our current understanding of the VTS molecules is necessarily incomplete. 
%
Note that, on the other hand, if our predictions are experimentally confirmed, it does not imply that our understanding of the VTS is complete.\\
%
\ankit{Added this as a template. Will change it soon.}
\begin{example}
%
In figure~\ref{fig:M1}, we present a VTS that has 3 compartments and 8 molecules, and a corresponding pairing matrix.
%
Molecules are numbered 0-7.
%
In the VTS, labels are a string of molecule ids, and an overline over an id indicates that the molecule is active.
%
Every molecule on the node is active.
%
The activity of the molecules on an edge are controlled
by presence of the other molecules on the edge.
%
On the right side of the figure, we show the pairing matrix.
%
An entry 1 represents pairing between molecules.
%
$\times$ represents no pairing.
%
Rows corresponds to the labels of edges, and
columns corresponds to the labels of nodes.
%
Every molecule flows on a cycle, ensuring steady state.
%
This is a 3-connected graph.
\end{example}
%
\input{pm-1-matrix}
%\subsection{Modeling and Symbolic Analysis of VTS: An Overview}
\label{subsec:graphmodel}

\subsection{Combinatorial problem and use of Formal methods}
\ankit{Needs revision; overlapping writing.}
The analysis of vesicle traffic systems is a difficult problem
because of the combinatorial scaling of possible traffic topologies and regulatory rules. 
%
%For example, we might want to check some conjecture of interest for all graphs of a certain structure. 
%
%If the analysis is done on a large number of graphs, the joined effect of large quantity of graphs and combinatorially many possible regulatory rules will be difficult to handle by any tool.
If the analysis is done on a large number of graphs, the joined effect of this with combinatorially many possible regulatory rules will be difficult to handle by any tool.
%
Precise analysis of the properties of the VTS would be hindered by the same reason.

@@

In contrast, here we seek understand properties of vesicle traffic networks over all possible update rules, not just for a sampled subset. 
%
We would also like to directly make statements about steady state properties, rather
than using transition dynamics to first search for steady states. 
%
Formal verification tools like model checkers, SAT and SMT solvers serve precisely this purpose. 
%
The tools model the computation symbolically for the graphs of finite size and can make prediction about the hypothesis for all possible states of the system without explicit enumeration of rules. 
%

% i.e., formally as a boolean function that given requisite labels of nodes and vesicles that returns true if the required regulations are met. 
%
%
% Note that the formulas define among other things constraints over
% paths between nodes in the graph model of VTS. Similarly, one can
% define fusion rules as constraints over the boolean functions modeling
% the regulations.  
\
% For example, the steady state condition, described informally earlier, can be defined by the formula shown in Listing~1.1.

%\begin{itemize}
%\item \srivas{Detailed BIO to Graph problem}
%\end{itemize}

%Given definitions of fusion and budding rules and the steady state conditions, whether a VTS meets maximum connectivity requirement, i.e. the LGC property, can be verified by checking if the formula show in Listing~1.2 is valid.  Here we have defined the property by checking over existence of any fusion/budding rules.  We can eliminate the existential quantifier if we are interested in checking the property for a particular fusion/budding rule.
%
%\subsection{Converting the Graph Problem into Boolean SAT problem}
%\label{subsec:satproblem}
%To convert the graph problem described in sec~\ref{subsec:graphmodeling}, into a boolean SAT problem, we need to define a scheme to represent graphs and boolean functions using a suitable set of propositional variables.  In our earlier work, we modeled the graph problem in C using arrays to model graphs and boolean function.  We then used the CBMC model checker to convert the graph problem into a SAT problem.  One of the challenges we had in our earlier work is dealing with quantifiers.  Note that the connectedness property defined in Listing~1.2 has quantifier alternation.  Even if we were to eliminate the existential quantifier by instantiating the problem for a fixed set of fusion rules, we would have embedded quantifiers in the antecedent of implications.   CBMC supports only a quantifier-free logic n its assertion language.  In our earlier work we used a combination of explicit enumeration at the C-level and clever use f non-determinism to eliminate alternation of quantifiers.  This enumeration was one source of bottleneck for scaling in our earlier work.  In the current work we eliminate this bottleneck by modeling the problem directly as a SAT instance using uninterpreted functions and recursive relations.  The details of te hnew encding will described in subsequent sections.

%A VTS is {\em $k$-connected} if every pair of compartments remain reachable after dropping $k-1$ vesicles.
%
%This property of VTSs has been considered informative and studied by~\cite{shukla}.
%
%Here we have build an {\em efficient} tool that studies properties of VTSs that are not $k$-connected, from some $k$. 
\subsection{Synthesis of VTS }
The current picture of the vesicle transport network is far from complete; we do not yet know how many of the cellular proteins reach their resident compartments within the cell, and new vesicle routes are being discovered every year ~\cite{nickel2018unconventional,weill2018toolbox}. 
%
Completing the vesicle traffic network is a difficult task for the many important reasons: 
\begin{enumerate}
	\item The core vesicle transport network is conserved across eukaryotes, the traffic network in different organisms \cite{richardson2015evolutionary,nishimoto2009differential,barlow2017seeing}, and even in different cell types within an organism can be different ~\cite{stoops2014trafficking,zhou2015arp2}.

@@

We have implemented the encoding in a flexible tool, which can handle a wide range of synthesis queries. 
%
We have applied our tool on several VTSs including
two found-in-nature VTSs.
%
Our experiments suggest that some of the synthesis problems are solvable by modern solvers and the synthesis technology may be useful for biological research. \\
%
%
\ankit{Added this as a template. Will change it soon.}
\begin{example}
	%
	In figure~\ref{fig:synth-vts}, we present a partial VTS that has 3 compartments and it's complete VTS by our tool. The synthesis is performed on the basis of steady state property and the  k-connectivity of the graph.

>>>>

The next step is membrane deformation, which involves cytosolic coat proteins. 
%
Coat proteins are recruited to the forming vesicle through their interaction with adaptor proteins. 
%
These are naturally curved proteins, and their assembly on membranes leads to membrane deformation.
%
A single kind of coat protein can bind to multiple kinds of adaptors, thereby allowing the packaging of multiple types of cargo within the same vesicle. 
%
There are three major types of coat proteins: Clathrin, COP2 and COP1. These act on different compartments: the clathrin coat operates at multiple compartments, and can interact with multiple adaptors proteins, whereas the COP2 and COP1 coat are only involved in vesicle formation at a single compartment: the ER and the Golgi respectively~\cite{faini2013vesicle}.
%
Finally, transport vesicles bud out from the source compartment~\cite{cocucci2014dynamin}.
%
The vesicle is now ready for transport and subsequent fusion with a target compartment.

@@

Target compartment recognition involves molecular switches called Rabs and  tethering complexes. Distinct compartments of the cell are associated with distinct Rabs. For example, Rab5 is associated early endosomes and  Rab7 with late endosomes\cite{rink2005rab}.  
%
Rabs are master regulators; in their active form Rabs recruit many downstream proteins, including tethering complexes, fusion regulators, motor proteins, etc, to their compartment ~\cite{muller2018molecular}.

Tethering complexes sequester vesicles to their target compartments by binding to the vesicle on one end and the target compartment on the other end, thereby bridging the two before fusion. 
%
They are also believed to function as chaperones for assembly of SNARE protein complexes, which enable vesicle fusion. 
%
A clear mechanism for action of these tethers is missing, but evidence from experiments indicate that tethers bind to vesicles through molecular interactions with vesicle coat proteins, vesicle, associated Rabs, SNAREs and SNARE regulators. 
%
They can also sense membrane curvature (vesicles are highly curved) ~\cite{baker2016chaperoning}.

The final step in vesicle transport is fusion of vesicle with the target compartment is brought about by SNARE proteins. 
%
There are four kinds of SNAREs: R, Qa, Qb, and Qc; and vesicle fusion requires formation of a SNARE complex containing one of each type. 
%
In all known cases, one SNARE protein of the complex is contributed by the vesicle and the other three SNAREs by the target compartment. 
%
A few SNARE proteins, such as SNAP25, contain both Qb- and Qc-SNARE motifs~\cite{yoon2018snare}.

%\somya{schematic of snare regulatory network}
\textbf{Schematic of snare regulatory network:}
In keeping with their central role in vesicle transport, both the activity and the localization of SNARE proteins are regulated. Notable regulators are Sec17/18 and the SM proteins.
%
Sec17/18 bring about disassembly of SNARE complexes post-fusion in order that these SNAREs may be reused. 
%
Whereas, SM proteins have following multiple modes of SNARE regulation: 
\begin{itemize}
	\item SM proteins can hold
	the Qa-SNAREs in an inhibited state and prevent or postpone their assembly into SNARE
	complexes.
	%
	\item Some SM proteins can act as a template upon which an early SNARE complex intermediate can
	form.
	%
	\item Finally, SM proteins bind the fully assembled
	SNARE complexes to protect them from premature disassembly by Sec17/18~\cite{baker2016chaperoning}.
	%
	\item SM proteins have also been shown to interact with tethers~\cite{yoon2018snare}.
\end{itemize}

@@

\subsection{The model}
\todo{Ankit: The Rothman model has no intuition; Add couple of lines relating it to VTS; Also, relate it to previous subsection. The connect is lacking.} The VTS model for this paper is inspired by~\cite{shukla2017discovering}. 
%
On the timescales of minutes, we have tried to capture important aspect of Rothman-Schekman-Sudhof (RSS) model~\cite{rothman2002machinery} of vesicle traffic system.
%
The following are the basic component and assumptions of our model. 
\begin{enumerate}
\item A cell is a set of compartments exchanging molecules via vesicles of uniform size.
\item On the timescale of minutes, compartments are neither created nor destroyed.
\item Each compartment is in steady state, gain and loss of the molecules is balance\todo{Ankit: Need intuition and also reasoning why this is okay to add this as an assumption}.
\item We assume that molecules are neither created nor destroyed.
\item Molecules move via vesicles of uniform size.
\item Identical vesicles have identical target compartments.
\item Fusion of vesicles to compartments is driven by specific SNARE pairing.
\item The identities of compartments and vesicles is encoded in their molecules, and identical vesicles fuse with identical target compartments.
\item The activity of a SNARE can be regulated by other molecules present on the same compartment
or vesicle.
\item An active SNARE pair is necessary and sufficient for fusion. 
\end{enumerate}

@@

\textbf{Steady state:} 
The vesicle transfer will change the molecular composition (distinct molecule count) on both the source compartment and the target compartment. 
%
In our model, we assume that the cell is in steady state. As a further simplification, we also assume that the incoming and outgoing flux is balanced for each of the molecular types at each compartment. 
Indeed, it is well accepted that on the scale of minutes and hours the molecular composition remains the same. 
%
In this paper, we refer to the steady state of the cell (referred to as \textit{homeostasis} in biological literature) as the stability condition over the annotated graph.

@@

%
%Properties of the VTS would be hindered by the same reason. 
%3-connected graphs and all possible variations of SNARE regulation rules. 
%
%The number of graphs of specified connectivity grows exponentially with the number of nodes: Table~\ref{tab:threec} shows how many 3-edge-connected graphs~\cite{a052448-oeis} exist (without parallel or self edges) as node number N increases.
\subsection{Graph connectivity is an interesting and accessible property of the VTS.} 
%\textbf{Properties of a VTS that satisfies all cell-biological constraints:} 
Interactions between the molecules of the VTS together produce a functional vesicle transport network, in which molecules flow between intracellular compartments using specified paths.
%
The molecules of the VTS, their interactions and modes of regulation have been worked out to a large extent through cell-biological experiments. 
%
But we do not know if this information is complete. In order to assess how completely we know the VTS, we need easily readable benchmarks.

@@

%\ankit{We need not specify Synthesis here. Only connectivity use next subsection to build this.}
%\somya{But this is the para where we explain why connectivity is interesting an biological property}
%\ankit{Yes; Both of our remark looks correct!}
The vesicle transport graphs available today for eukaryotic organisms are incomplete themselves. 
%
In this situation, we use our result as a tool to predict missing edges which satisfy the graph connectivity constraint. 
%
These predictions are intended to be used as guides for future experiments. 
%
A mismatch between our predictions and experiments would indicate that our current understanding of the VTS molecules is necessarily incomplete. 
%
Note that, on the other hand, if our predictions are experimentally confirmed, it does not imply that our understanding of the VTS is complete.\\
%

\begin{example}
%
In figure~\ref{fig:M1}, we present a VTS that has 3 compartments and 8 molecules, and a corresponding pairing matrix.
%
Molecules are numbered 0-7.
%
In the VTS, labels are a string of molecule ids, and an overline over an id indicates that the molecule is active.
%
Every molecule on the node is active.
%
The activity of the molecules on an edge are controlled
by presence of the other molecules on the edge.
%
On the right side of the figure, we show the pairing matrix.
%
An entry 1 represents pairing between molecules.
%
$\times$ represents no pairing.
%
Rows corresponds to the labels of edges, and
columns corresponds to the labels of nodes.
%
Every molecule flows on a cycle, ensuring steady state.
%
This is a 3-connected graph.
\end{example}
%
\input{pm-1-matrix}
%\subsection{Modeling and Symbolic Analysis of VTS: An Overview}
\label{subsec:graphmodel}

\subsection{Combinatorial problem and use of Formal methods}
The analysis of vesicle traffic systems is a difficult problem
because of the combinatorial scaling of possible traffic topologies and regulatory rules. 
%
%For example, we might want to check some conjecture of interest for all graphs of a certain structure. 
%
%If the analysis is done on a large number of graphs, the joined effect of large quantity of graphs and combinatorially many possible regulatory rules will be difficult to handle by any tool.
If the analysis is done on a large number of graphs, the joined effect of this with combinatorially many possible regulatory rules will be difficult to handle by any tool.
%
Precise analysis of the properties of the VTS would be hindered by the same reason.

@@

In contrast, here we seek understand properties of vesicle traffic networks over all possible update rules, not just for a sampled subset. 
We would also like to directly make statements about steady state properties, rather
than using transition dynamics to first search for steady states. 
Formal verification tools like model checkers, SAT and SMT solvers serve precisely this purpose. 
The tools model the computation symbolically for the graphs of finite size and can make prediction about the hypothesis for all possible states of the system without explicit enumeration of rules. 
% i.e., formally as a boolean function that given requisite labels of nodes and vesicles that returns true if the required regulations are met. 
% Note that the formulas define among other things constraints over
% paths between nodes in the graph model of VTS. Similarly, one can
% define fusion rules as constraints over the boolean functions modeling
% the regulations.  
% For example, the steady state condition, described informally earlier, can be defined by the formula shown in Listing~1.1.
%\begin{itemize}
%\item \srivas{Detailed BIO to Graph problem}
%\end{itemize}
%Given definitions of fusion and budding rules and the steady state conditions, whether a VTS meets maximum connectivity requirement, i.e. the LGC property, can be verified by checking if the formula show in Listing~1.2 is valid.  Here we have defined the property by checking over existence of any fusion/budding rules.  We can eliminate the existential quantifier if we are interested in checking the property for a particular fusion/budding rule.
%\subsection{Converting the Graph Problem into Boolean SAT problem}
%\label{subsec:satproblem}
%To convert the graph problem described in sec~\ref{subsec:graphmodeling}, into a boolean SAT problem, we need to define a scheme to represent graphs and boolean functions using a suitable set of propositional variables.  In our earlier work, we modeled the graph problem in C using arrays to model graphs and boolean function.  We then used the CBMC model checker to convert the graph problem into a SAT problem.  One of the challenges we had in our earlier work is dealing with quantifiers.  Note that the connectedness property defined in Listing~1.2 has quantifier alternation.  Even if we were to eliminate the existential quantifier by instantiating the problem for a fixed set of fusion rules, we would have embedded quantifiers in the antecedent of implications.   CBMC supports only a quantifier-free logic n its assertion language.  In our earlier work we used a combination of explicit enumeration at the C-level and clever use f non-determinism to eliminate alternation of quantifiers.  This enumeration was one source of bottleneck for scaling in our earlier work.  In the current work we eliminate this bottleneck by modeling the problem directly as a SAT instance using uninterpreted functions and recursive relations.  The details of te hnew encding will described in subsequent sections.
%A VTS is {\em $k$-connected} if every pair of compartments remain reachable after dropping $k-1$ vesicles.
%This property of VTSs has been considered informative and studied by~\cite{shukla}.
%Here we have build an {\em efficient} tool that studies properties of VTSs that are not $k$-connected, from some $k$. 
\subsection{Synthesis of VTS }
The current picture of the vesicle transport network is far from complete; we do not yet know how many of the cellular proteins reach their resident compartments within the cell, and new vesicle routes are being discovered every year ~\cite{nickel2018unconventional,weill2018toolbox}. 
Completing the vesicle traffic network is a difficult task for the many important reasons: 
\begin{enumerate}
	\item The core vesicle transport network is conserved across eukaryotes, the traffic network in different organisms \cite{richardson2015evolutionary,nishimoto2009differential,barlow2017seeing}, and even in different cell types within an organism can be different ~\cite{stoops2014trafficking,zhou2015arp2}.

@@

We have implemented the encoding in a flexible tool, which can handle a wide range of synthesis queries. 
%
We have applied our tool on several VTSs including
two found-in-nature VTSs.
%
Our experiments suggest that some of the synthesis problems are solvable by modern solvers and the synthesis technology may be useful for biological research.
%
\begin{example}
	%
	In figure~\ref{fig:synth-vts}, we present a partial VTS that has 3 compartments and it's complete VTS by our tool. The synthesis is performed on the basis of steady state property and the  k-connectivity of the graph.
